---
title: "Lecture 5: Expectation and Variance"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---

<!-- 
Before running the document, make sure `learnr` package is available. 
If missing, you can install the package with `install.packages('learnr')`.
-->

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## R Preparation

### Annotations in `ggplot`

Previously, we used `ggplot` to plot curves and points. 

```{r echo=TRUE}
library(ggplot2)
x <- 1:10
y <- sqrt(x)
base_plot <- ggplot(mapping = aes(x = x, y = y)) +
  theme_classic() +
  geom_point() +
  geom_line() +
  labs(title = "y is square root of x", x = "x", y = "square root of x")
base_plot
```

You can add annotations using `annotate()`. Annotations can be in different forms. 

+  `text`

```{r echo=TRUE}
base_plot +  # you can save a ggplot to a variable and keep adding layers
  annotate(
    "text",  # specify the type of annotation
    x = 5, y = sqrt(5), # specify reference position of the text
    label = "Tell me something   ", # text to annotate
    hjust = 1, vjust = 0    # you can adjust the justification of text
  )
```

+  `point`

```{r echo=TRUE}
base_plot +  
  annotate(
    "point",  # specify the type of annotation
    x = 5, y = sqrt(5), # specify position 
    shape = 17, color = "red", size = 7
  )
```

+  `segment`

```{r echo=TRUE}
base_plot +  
  annotate(
    "segment",  # specify the type of annotation
    x = 5, y = sqrt(5), # specify position of text
    xend = 0, yend = sqrt(5), # a segment also requires ends
    color = "red", linetype = "dotted"
  )
```

For more details, check `?annoate`.

```{r helpannoatae, exercise=TRUE}
?annotate
```

You can also draw horizontal and vertical lines by specifying the y and x intercept values only using `geom_hline()` and `geom_vline()` respectively.

```{r echo=TRUE}
base_plot +
  geom_hline(yintercept = 2, color = "blue") +
  geom_vline(xintercept = 4, color = "darkgreen")
```


#### Practice

In the code chunk below, `base_plot` from above examples is available. Practice adding different annotations.

```{r annotate-practice-setup}
library(ggplot2)
x <- 1:10
y <- sqrt(x)
base_plot <- ggplot(mapping = aes(x = x, y = y)) +
  theme_classic() +
  geom_point() +
  geom_line() +
  labs(title = "y is square root of x", x = "x", y = "square root of x")
```

```{r annotate-practice, exercise=TRUE, exercise.lines=20}
base_plot
```

```{r annotate-practice-hint}
base_plot +
  annotate("text", x = 10, y = sqrt(10), label = "f(x) = sqrt(x)",
           vjust = -0.5, hjust = 1) +
  geom_vline(xintercept = 9 / 4, linetype = "dotted") +
  geom_hline(yintercept = 3 / 2, linetype = "dotted") +
  annotate("segment", x = 0, y = 3 / 2, xend = 9 / 4, yend = 3 / 2, 
           color = "pink", linewidth = 2) +
  annotate("segment", x = 9 /4, y = 1, xend = 9 / 4, yend = 3 / 2, 
           color = "pink", linewidth = 2) +
  annotate("point", x = 9 / 4, y = 3 / 2, 
           shape = 19, size = 7, color = "pink")
```

## Estimating Expectations via Simulation

Recall, we have been estimating probabilities via simulation based on the perspective that a probability is a long-term relative frequency. We can also show that a probability of an event is an expectation of a Bernoulli random variable. You will show this in the next Weekly Activity. In turn, it tells us that we have been estimating expectations via simulations. 

### Example: My Coffee Shop

Consider $D\sim\text{Pois}(4)$ from the lecture. Recall that we derived $E[D]=4$. Let's verify that the $E[D]$ represents the centre of probability mass and that we can retract the value via simulation.

1. Generate $100,000$ simulated values of $\text{Pois}(4)$.
2. Verify that $E[D]$ represents the centre of probability mass.
  a. Sum distances from $E[D]$ to the simulated values that are less than or equal to $E[D]$.
  b. Sum distances from $E[D]$ to the simulated values that are greater than $E[D]$.
  c. Compare the sums.
3. Compute the average of the values.
  a. Verify the average value is close to $E[D]$.
  


```{r coffee-shop, exercise=TRUE, exercise.lines=10}

```


```{r coffee-shop-hint-1}
N <- 10^5
lambda <- 4
sims <- rpois(N, lambda)
sim_mean <- sum(sims) / N  # average
```

```{r coffee-shop-hint-2}
N <- 10^5
lambda <- 4
sims <- rpois(N, lambda)
sim_mean <- mean(sims)     # mean() is the function for computing the average
```

```{r coffee-shop-solution}
N <- 10^5
lambda <- 4
sims <- rpois(N, lambda)
# probability masses for values less than or equal to E[D]
sum(lambda - sims[sims <= lambda])
# probability masses for values greater than E[D]
sum(sims[sims > lambda] - lambda)
# average of the values
sim_mean <- mean(sims)
```

