---
title: "Lecture 5: Expectation and Variance"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---

<!-- 
Before running the document, make sure `learnr` package is available. 
If missing, you can install the package with `install.packages('learnr')`.
-->

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
library(gganimate)
knitr::opts_chunk$set(echo = FALSE)
```


## Computing Expectation and Variance

Consider $N\sim\text{Bin}(20, 0.4)$. Below is the pmf of $N$.

```{r echo=FALSE}
distN <- tibble(x = 0:20) %>%
  mutate(d = dbinom(x, 20, 0.4))
ggplot(distN, aes(x = x, y = d)) +
  theme_classic() +
  geom_bar(stat = "identity", width = 0.05) +
  geom_point() +
  labs(title = "Probability mass function of N, p(n)",
       y = "p", x = "n")
```

### Compute $E(N)$ in R

```{r nmean, exercise=TRUE, exercise.lines=5}
x <- 0:20 # all possible values

```

### Compute $\text{Var}(N)$ in R

```{r nvar, exercise=TRUE, exercise.lines=5}
x <- 0:20 # all possible values

```

## Note on `dgeom` in R

In class, we defined the geometric random variable to represent the number of trials until success including the successful trial with pmf $p$.

$$p(x)=p(1-p)^{x-1}$$

In R, `?dgeom` opens the help page for the geometric distribution functions. It reveals that R uses a different pmf. Namely,

$$p(x)=p(1-p)^x.$$

R takes alternative definition of a geometric random variable where it represents the number of failures **before** the first success. 

### Working with R's Representation

Suppose $Y$ represents the number of trials _until_ the first success. There is a 0.3 probability of success for each trial. Compute and plot the pmf $p(y)$ and the cdf $F(y)$ for $y=1,2,\ldots,20$, respectively.

```{r dgeom, exercise=TRUE, exercise.lines=10}
y <- 1:20 # values of interest

```